---
title: "plot_testing"
author: "Seth (The True) Villanueva"
date: "2023-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(terra)
# library(tidyterra)        # for mapping terra rasters
library(sf)
library(tmap)
library(mapproj)
library(maptools)
library(ggnewscale)
library(here)
```


### define global variables and functions

```{r}
data_path <- 'fish_data_app/data'

source(here(data_path, 'helper_fxns.R'))
```


### read in macro datasets

```{r}
fish_info <- read_csv(here(data_path, 'fish_info.csv'))

rgns_geog <- read_csv(here(data_path, 'spatial', 'meow_rgns.csv'))
rgn_names_ohi <- read_csv(here(data_path, 'spatial', 'rgn_names_ohi.csv'))
rgns_geog_raster <- raster(here(data_path, 'spatial', 'meow_rgns_mol.tif'))
```


```{r}
species_choice <- 'thunnus albacares'

top_stress <- fish_info %>%
      filter(species %in% species_choice) %>%
      select(stressor,vuln) %>%
      arrange(desc(vuln)) %>% 
      filter(vuln == max(vuln)) %>% 
      select(stressor) %>% 
      mutate(stressor = str_replace(stressor, pattern = "_", replacement = " ")) %>% 
      as.list()


lengths(top_stress) == 2

```




# Make map of environmental stressors and species range

Where are I/Os located?

Inputs:

  Region
    realm - data/spatial/
                  meow_rgns.csv
                      $realm_code
                      $realm
                  meow_rgns_mol.tif             # maps of the realms?
  Stressor
    name - data/fish_info
                  $species
    area - data/stressor_maps
  Species
    name - data/fish_info
                  $stressor
    range - data/species_ranges
                  `data_source'_spp_mol_`species_name_or_number'.csv
    vulnerability - data/fish_info
                            $vuln

Outputs:
  

### make species lists for conditionals

```{r}
am_species <- c('chanos chanos', 
                'gadus morhua', 
                'mallotus villosus',
                'oncorhynchus mykiss',
                'salmo salar',
                'trichiurus lepturus')

iucn_species <- c('brevoortia patronus',
                  'clupea harengus',
                  'engraulis japonicus',
                  'engraulis ringens',
                  'katsuwonus pelamis',
                  'sardina pilchardus',
                  'sardinella longiceps',
                  'scomber japonicus',
                  'scomber scombrus',
                  'thunnus albacares')

iucn_species_dict <- c('brevoortia patronus' = '191208',
                  'clupea harengus' = '155123',
                  'engraulis japonicus' = '170306',
                  'engraulis ringens' = '183775',
                  'katsuwonus pelamis' = '170310',
                  'sardina pilchardus' = '198580',
                  'sardinella longiceps' = '154989',
                  'scomber japonicus' = '98969433',
                  'scomber scombrus' = '170354',
                  'thunnus albacares' = '21857')

# make dictionary with all stressor and file names
stressor_tif_dict <- c("air_temp" = "",   
                        "biomass_removal" = "",
                        "bycatch" = "bycatch_benthic_2017",  # there are two bycatch files with different suffixes?
                        "entanglement_macroplastic" = "",
                        "eutrophication_nutrient_pollution" = "nutrient_2020",
                        "habitat_loss_degradation" = "",
                        "inorganic_pollution" = "",
                        "light_pollution" = "light_2018",
                        "marine_heat_waves" = "sst_extremes_2020",             # is this where sst_extreme should go?
                        "noise_pollution" = "",
                        "ocean_acidification" = "ocean_acidification_2020",
                        "oceanographic" = "",
                        "organic_pollution" = "",
                        "plastic_pollution_microplastic" = "microplastics_2015",
                        "poisons_toxins" = "",
                        "salinity" = "",
                        "sedimentation" = "",
                        "sst_rise" = "spp_max_temp",     # this is located in a different location than the others, and is only part of the file name
                        "storm_disturbance" = "",
                        "uv_radiation" = "uv_radiation_2020",
                        "wildlife_strike" = ""
)
# usused maps: benth, bycatch_pelagic, direct_human, sea_level_rise, shipping

```

### Make region limits

### Extract stressor and species specific rasters

```{r, message = FALSE}

region_choice <- 1
stressor_choice <- c('uv_radiation')
species_choice <- 'oncorhynchus mykiss'

# format inputs for feeding to file chains
species_choice_formatted <- sub(' ', '_', tolower(species_choice))
species_name_file <- species_choice_formatted

# set source depending on species
src <- ''
if (species_choice %in% am_species) {
  src <- 'am'
}
if (species_choice %in% iucn_species) {
  src <- 'iucn'
  species_name_file <- iucn_species_dict[species_choice]
}



##### format STRESSOR file names for calls according to choice

# initialize path and name variables
stressor_tif_name <- ''
stressor_tif_path_addition <- ''
stressor_tif_path <- ''

# add to path and name for sea surface temperature maps
if (stressor_choice == "sst_rise") {
  
  sst_tif_prefix <- '_spp_max_temp_'
  stressor_tif_path_addition <- 'sst_rise_maps'
  
  stressor_tif_name <- paste(src, sst_tif_prefix, species_name_file, sep = '')
  stressor_tif_path_addition <- 'sst_rise_maps'
  
} else {                                    # for all other files, refer to the dictionary
  
  stressor_tif_name <- stressor_tif_dict[stressor_choice]
  stressor_tif_path_addition <- 'stressor_maps'

}

stressor_tif_folder <- paste(data_path, '/', stressor_tif_path_addition, sep = '')
stressor_tif_path <- paste(stressor_tif_folder, '/', stressor_tif_name, '.tif', sep = '')



##### Format SPECIES range file name for calls

species_range_file <- paste(src, '_spp_mol_', species_name_file, sep = '')
species_range_csv_path <- paste(data_path, '/species_ranges/', species_range_file, '.csv', sep = '')

species_range_df = read_csv(here(species_range_csv_path))

# csv processing depends on the data source
if (src == 'iucn') {
  species_which <- 'presence'
} else {          # for src == 'am'
  species_range_df <- species_range_df %>% 
    filter(prob >= 0.5) %>% 
    drop_na()
  species_which <- 'prob'
}



##### Capture species vulernability to the chosen stressor

species_vuln <- fish_info %>% 
  filter(species == species_choice, stressor == stressor_choice) %>% 
  pull(vuln)



##### Generate rasters for stressor and species range
# Also change the CRS
crs_proj <- 'epsg:4326'

stressor_rast <- rast(here(stressor_tif_path)) %>% 
  project(crs_proj)



# call helper function to make raster from csv
species_rast <- map_to_mol(species_range_df,
                           by = 'cell_id',
                           which = species_which,
                           ocean_mask = TRUE) %>% 
  project(crs_proj)  


# Make separate rasters that are mutually exclusive; for species stress and stressor
stressor_intersect <- terra::mask(stressor_rast, species_rast)      # crop the stressor map to where the species range is
product_rast <- stressor_intersect * species_rast                   # calculate species stress
inverse_product_rast <- terra::mask(stressor_rast, product_rast, inverse=TRUE)    # make map of stressor that excludes species stress


```


### make plot

We want to have a base color opacity for the species range equal to the vulnerability, then when it overlaps with stressor the color should scale according to the product of vulernability and stressor amount

It might be best to make a series of raster files to reflect the color options we want. For example, for 1 species, we would need 2 rasters: 1 for stressor area EXCLUDING stressor overlap with species range, then 1 for species range INCLUDING overlap with stressor area. The species range should then be able to range in color from a minimum that reflects the base vulernability and the scaling with stressor exposure.

First attempt with ggplot (can later maybe attempt with tmap)
```{r}
# species_stress_df_var <- 'stress'
species_stress_df <- as.data.frame(x = product_rast, xy = TRUE) %>%
  rename('species_stress' = 'uv_radiation_2020')
#   rename(!!species_stress_df_var := stressor_tif_name)                # NEED TO GET THIS GENERAL FUNCTION WORKING
stressor_df <- as.data.frame(x = inverse_product_rast, xy = TRUE) %>%
  rename('stress' = 'uv_radiation_2020')

species_stress_map <- ggplot() +
  geom_tile(data = species_stress_df, aes(x = x, y = y, fill = species_stress)) +
  # scale_color_viridis_c() +
  scale_fill_gradient(low = 'orange1', high = 'red4') +
  new_scale_fill() +
  geom_tile(data = stressor_df, aes(x = x, y = y, fill = stress)) +
  scale_fill_gradient(low = 'cyan1', high = 'blue4') +
  #scale_color_distiller('stress', palette = "Spectral") +
  # coord_sf(expand = 0) +
  # coord_cartesian(xlim = c(-13000000, -8000000), ylim = c(0, 10000000)) +
  theme_void()

species_stress_map

```

### Recentering the map to the data
Adapted from: https://stackoverflow.com/questions/10620862/use-different-center-than-the-prime-meridian-in-plotting-a-world-map/10629246#10629246
```{r}
library(sp)
library(maps)
library(maptools)   ## map2SpatialLines(), pruneMap()
library(rgdal)      ## CRS(), spTransform()
library(rgeos)      ## readWKT(), gIntersects(), gBuffer(), gDifference() 

## Convert a "maps" map to a "SpatialLines" map
makeSLmap <- function() {
    llCRS <- CRS("+proj=longlat +ellps=WGS84")
    wrld <- map("world", interior = FALSE, plot=FALSE,
            xlim = c(-179, 179), ylim = c(-89, 89))
    wrld_p <- pruneMap(wrld, xlim = c(-179, 179))
    map2SpatialLines(wrld_p, proj4string = llCRS)
}

## Clip SpatialLines neatly along the antipodal meridian
sliceAtAntipodes <- function(SLmap, lon_0) {
    ## Preliminaries
    long_180 <- (lon_0 %% 360) - 180
    llCRS  <- CRS("+proj=longlat +ellps=WGS84")  ## CRS of 'maps' objects
    eqcCRS <- CRS("+proj=eqc")
    ## Reproject the map into Equidistant Cylindrical/Plate Caree projection 
    SLmap <- spTransform(SLmap, eqcCRS)
    ## Make a narrow SpatialPolygon along the meridian opposite lon_0
    L  <- Lines(Line(cbind(long_180, c(-89, 89))), ID="cutter")
    SL <- SpatialLines(list(L), proj4string = llCRS)
    SP <- gBuffer(spTransform(SL, eqcCRS), 10, byid = TRUE)
    ## Use it to clip any SpatialLines segments that it crosses
    ii <- which(gIntersects(SLmap, SP, byid=TRUE))
    # Replace offending lines with split versions
    # (but skip when there are no intersections (as, e.g., when lon_0 = 0))
    if(length(ii)) { 
        SPii <- gDifference(SLmap[ii], SP, byid=TRUE)
        SLmap <- rbind(SLmap[-ii], SPii)  
    }
    return(SLmap)
}

## re-center, and clean up remaining streaks
recenterAndClean <- function(SLmap, lon_0) {
    llCRS <- CRS("+proj=longlat +ellps=WGS84")  ## map package's CRS
    newCRS <- CRS(paste("+proj=eqc +lon_0=", lon_0, sep=""))
    ## Recenter 
    SLmap <- spTransform(SLmap, newCRS)
    ## identify remaining 'scratch-lines' by searching for lines that
    ## cross 2 of 3 lines of longitude, spaced 120 degrees apart
    v1 <-spTransform(readWKT("LINESTRING(-62 -89, -62 89)", p4s=llCRS), newCRS)
    v2 <-spTransform(readWKT("LINESTRING(58 -89, 58 89)",   p4s=llCRS), newCRS)
    v3 <-spTransform(readWKT("LINESTRING(178 -89, 178 89)", p4s=llCRS), newCRS)
    ii <- which((gIntersects(v1, SLmap, byid=TRUE) +
                 gIntersects(v2, SLmap, byid=TRUE) +
                 gIntersects(v3, SLmap, byid=TRUE)) >= 2)
    SLmap[-ii]
}

## Put it all together:
Recenter <- function(lon_0 = -100, grid=FALSE, ...) {                        
    SLmap <- makeSLmap()
    SLmap2 <- sliceAtAntipodes(SLmap, lon_0)
    recenterAndClean(SLmap2, lon_0)
}

## Try it out
par(mfrow=c(2,2), mar=rep(1, 4))
plot(Recenter(-90), col="grey40"); box() ## Centered on 90w 
plot(Recenter(0),   col="grey40"); box() ## Centered on prime meridian
plot(Recenter(90),  col="grey40"); box() ## Centered on 90e
plot(Recenter(180), col="grey40"); box() ## Centered on International Date Line
```







```{r}
# Attempting to use sf instead of geom_raster

# species_stress_poly <- terra::as.polygons(product_rast) %>% 
#   st_as_sf()
# 
# stress_poly <- terra::as.polygons(product_rast) %>% 
#   st_as_sf()
# 
# species_stress_map <- ggplot() + 
#   geom_sf(data = species_stress_poly, fill = 'grey30', color = 'black') +
#   # scale_color_viridis_c() +
#   scale_fill_gradient(low = 'orange1', high = 'red4') +
#   new_scale_fill() +
#   geom_raster(data = stressor_df, aes(x = x, y = y, fill = stress)) + 
#   scale_fill_gradient(low = 'cyan1', high = 'blue4') +
#   #scale_color_distiller('stress', palette = "Spectral") +
#   coord_sf(expand = 0) +
#   theme_void()
# 
# species_stress_map
```

```{r}
# Attempting to make stressor map with species range polygons overlayed
# need a way of making fill opaque

stressor_df <- as.data.frame(x = stressor_rast, xy = TRUE) %>%
  rename('stress' = 'uv_radiation_2020')

species_poly <- terra::as.polygons(species_rast) %>%
  st_as_sf()

species_stress_map <- ggplot() +
  geom_raster(data = stressor_df, aes(x = x, y = y, fill = stress)) +
  scale_fill_gradient(low = 'cyan1', high = 'blue4') +
  #scale_color_distiller('stress', palette = "Spectral") +
  new_scale_fill() +
  geom_sf(data = species_stress_poly, color = 'black') +
  # scale_color_viridis_c() +
  coord_sf(expand = 0) +
  theme_void()

species_stress_map
```


Because of overlap, we may only be able to show 1 stressor and 1 species; unless species dont overlap

Another idea: Make two versions of stressor where one is the intersect with species range and one is the opposite, then just do the OUTLINE of the species ranges. This way, only the region where stressor and species overlap require a different color






Things to ask casey:

1. is the method I'm using to subset the data the best way of doing it for spatial data?
2. Am I missing something, since there are more stressors than maps?
3. How to treat similar maps like shipping large and small; combine somehow?
      shipping large for wildlife strike
4. Where do realm bounds come from?

5. best way of operating on a list of fish species
      And best way of adding map raster areas depending on a variable number of species





drop stressors for which all fish have a vuln of 0





maybe do a single species and then allow multiple stressors. Then we can show the total impact to the species as the sum of the stressors






