---
title: "plot_testing"
author: "Seth (The True) Villanueva"
date: "2023-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(terra)
library(sf)
library(tmap)
library(Dict)
library(here)
```


### define global variables and functions

```{r}
data_path <- 'fish_data_app/data'

source(here(data_path, 'helper_fxns.R'))
```


### read in macro datasets

```{r}
fish_info <- read_csv(here(data_path, 'fish_info.csv'))

rgns_geog <- read_csv(here(data_path, 'spatial', 'meow_rgns.csv'))
rgn_names_ohi <- read_csv(here(data_path, 'spatial', 'rgn_names_ohi.csv'))
rgns_geog_raster <- raster(here(data_path, 'spatial', 'meow_rgns_mol.tif'))
rgns_geog_raster
plot(rgns_geog_raster)
```

```{r}
eez_raster <- raster(here(data_path, 'spatial', 'eez_mol.tif'))
eez_raster
plot(eez_raster)

ocean_area_raster <- raster(here(data_path, 'spatial', 'ocean_area_mol.tif'))
ocean_area_raster
plot(ocean_area_raster)
```





# Make map of environmental stressors and species range

Where are I/Os located?

Inputs:

  Region
    realm - data/spatial/
                  meow_rgns.csv
                      $realm_code
                      $realm
                  meow_rgns_mol.tif             # maps of the realms?
  Stressor
    name - data/fish_info
                  $species
    area - data/stressor_maps
  Species
    name - data/fish_info
                  $stressor
    range - data/species_ranges
                  `data_source'_spp_mol_`species_name_or_number'.csv
    vulnerability - data/fish_info
                            $vuln

Outputs:
  

### make species lists for conditionals

```{r}
am_species <- c('chanos chanos', 
                'gadus morhua', 
                'mallotus villosus',
                'oncorhynchus mykiss',
                'salmo salar',
                'trichiurus lepturus')

iucn_species <- c('brevoortia patronus',
                  'clupea harengus',
                  'engraulis japonicus',
                  'engraulis ringens',
                  'katsuwonus pelamis',
                  'sardina pilchardus',
                  'sardinella longiceps',
                  'scomber japonicus',
                  'scomber scombrus',
                  'thunnus albacares')

iucn_species_dict <- c('brevoortia patronus' = '191208',
                  'clupea harengus' = '155123',
                  'engraulis japonicus' = '170306',
                  'engraulis ringens' = '183775',
                  'katsuwonus pelamis' = '170310',
                  'sardina pilchardus' = '198580',
                  'sardinella longiceps' = '154989',
                  'scomber japonicus' = '98969433',
                  'scomber scombrus' = '170354',
                  'thunnus albacares' = '21857')

# make dictionary with all stressor and file names
stressor_tif_dict <- c("air_temp" = "sst_extremes_2020",   # is this where sst_extreme should go?
                        "biomass_removal" = "",
                        "bycatch" = "bycatch_benthic_2017",  # there are two bycatch files with different suffixes?
                        "entanglement_macroplastic" = "",
                        "eutrophication_nutrient_pollution" = "nutrient_2020",
                        "habitat_loss_degradation" = "",
                        "inorganic_pollution" = "",
                        "light_pollution" = "light_2018",
                        "marine_heat_waves" = "",
                        "noise_pollution" = "",
                        "ocean_acidification" = "ocean_acidification_2020",
                        "oceanographic" = "",
                        "organic_pollution" = "",
                        "plastic_pollution_microplastic" = "microplastics_2015",
                        "poisons_toxins" = "",
                        "salinity" = "",
                        "sedimentation" = "",
                        "sst_rise" = "spp_max_temp",     # this is located in a different location than the others, and is only part of the file name
                        "storm_disturbance" = "",
                        "uv_radiation" = "uv_radiation_2020",
                        "wildlife_strike" = ""
)
# usused maps: benth, bycatch_pelagic, direct_human, sea_level_rise, shipping

```


### Extract stressor and species specific rasters

```{r, message = FALSE}

region_choice <- 1
stressor_choice <- 'uv_radiation'
species_choice <- c('oncorhynchus mykiss')

# format inputs for feeding to file chains
species_choice_formatted <- sub(' ', '_', tolower(species_choice))
species_name_file <- species_choice_formatted

# set source depending on species
src <- ''
if (species_choice %in% am_species) {
  src <- 'am'
}
if (species_choice %in% iucn_species) {
  src <- 'iucn'
  species_name_file <- iucn_species_dict[species_choice]
}



##### format STRESSOR file names for calls according to choice

# initialize path and name variables
stressor_tif_name <- ''
stressor_tif_path_addition <- ''
stressor_tif_path <- ''

# add to path and name for sea surface temperature maps
if (stressor_choice == "sst_rise") {
  
  sst_tif_prefix <- '_spp_max_temp_'
  stressor_tif_path_addition <- 'sst_rise_maps'
  
  stressor_tif_name <- paste(src, sst_tif_prefix, species_name_file, sep = '')
  stressor_tif_path_addition <- 'sst_rise_maps'
  
} else {                                    # for all other files, refer to the dictionary
  
  stressor_tif_name <- stressor_tif_dict[stressor_choice]
  stressor_tif_path_addition <- 'stressor_maps'

}

stressor_tif_folder <- paste(data_path, '/', stressor_tif_path_addition, sep = '')
stressor_tif_path <- paste(stressor_tif_folder, '/', stressor_tif_name, '.tif', sep = '')



##### Format SPECIES range file name for calls

species_range_file <- paste(src, '_spp_mol_', species_name_file, sep = '')
species_range_csv_path <- paste(data_path, '/species_ranges/', species_range_file, '.csv', sep = '')

species_range_df = read_csv(here(species_range_csv_path))

# csv processing depends on the data source
if (src == 'iucn') {
  species_which <- 'presence'
} else {          # for src == 'am'
  species_range_df <- species_range_df %>% 
    filter(prob >= 0.5)
  species_which <- 'prob'
}



##### Capture species vulernability to the chosen stressor

species_vuln <- fish_info %>% 
  filter(species == species_choice, stressor == stressor_choice) %>% 
  pull(vuln)



##### Generate rasters for stressor and species range

stressor_rast <- raster(here(stressor_tif_path))
# call helper function to make raster from csv
# species_rast <- map_to_mol(species_range_df,
#                            by = 'cell_id',
#                            which = species_which,
#                            xfm == NULL,
#                            ocean_mask = TRUE)


plot(stressor_rast)

```


### make plot

We want to have a base color opacity for the species range equal to the vulnerability, then when it overlaps with stressor the color should scale according to the product of vulernability and stressor amount

It might be best to make a series of raster files to reflect the color options we want. For example, for 1 species, we would need 2 rasters: 1 for stressor area EXCLUDING stressor overlap with species range, then 1 for species range INCLUDING overlap with stressor area. The species range should then be able to range in color from a minimum that reflects the base vulernability and the scaling with stressor exposure.

```{r}

```







Things to ask casey:

1. is the method I'm using to subset the data the best way of doing it for spatial data?
2. Am I missing something, since there are more stressors than maps?
3. How to treat similar maps like shipping large and small; combine somehow?
      shipping large for wildlife strike
4. Where do realm bounds come from?

5. best way of operating on a list of fish species
      And best way of adding map raster areas depending on a variable number of species





drop stressors for which all fish have a vuln of 0





maybe do a single species and then allow multiple stressors. Then we can show the total impact to the species as the sum of the stressors






