---
title: "plot_testing"
author: "Seth (The True) Villanueva"
date: "2023-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(tmap)
library(here)
```


### fish info

```{r}
fish_info <- read_csv(here('fish_data_app/data', 'fish_info.csv'))
```



# Make map of environmental stressors and species range

Where are I/Os located?

Inputs:

  Region
    realm - data/spatial/
                  meow_rgns.csv
                      $realm_code
                      $realm
                  meow_rgns_mol.tif             # maps of the realms?
  Stressor
    name - data/fish_info
                  $species
    area - data/stressor_maps
  Species
    name - data/fish_info
                  $stressor
    range - data/species_ranges
                  `data_source'_spp_mol_`species_name_or_number'.csv
    vulnerability - data/fish_info
                            $vuln

Outputs:
  

### subset data

```{r}
region_choice <- '01'
stressor_choice <- 'air_temp'
species_choice <- c('Chanos Chanos')

# format inputs for feeding to file chains
species_choice_formatted <- sub(' ', '_', tolower(species_choice))


```


### make plot

```{r}

```












